### 2009-12-14 docs/marcello-dissertation/thesis/chapter10-appendix.tex
Final Revision After Prof. Puder's Audio Feedback. Chapters 7 and Chapter 8 were included together. Chapter 10 also added as the appendix, instead of chapter.
* Chapter 7 

- The mouse experiment was used as demo purpose and to show the DSP use of the persistence. The data used was the for the workload used the YSI Generator from Teresa, which generated the SAME values each time… In this way, it was NOT random… I added the proper random functions in everything, including the parameter passed to the method to generate a given number of generated values. 

- The workload uses the amount of data

NetBEAMS uses the concept of the modularization through the use of the
Producer-Consumer paradigm described in the previous section. Since the DSP
Components are essentially OSGi bundles, the interoperability between
them is given by the OSGi Framework, described by an artifact called
MANIFEST.MF (Listing \ref{file:OSGi-manifest}). In general, an OSGi bundle must
provide specifications that describe the module to be published into the OSGi
Framework. In this way, the main properties of the OSGi MANIFEST.MF artifact
can used by the DSP Data Persistence can be summarized as follows:

\begin{itemize}
  \item \textbf{Bundle-Activator}: The name of the instance of an OSGi
  Activator class, responsible to manage the bundle. The implemented class
  ``DataPersistenceActivator'' provides the activation mechanisms for the
  services of the DSP Data Persistence Component;
  \item \textbf{Bundle-ClassPath}: the necessary Java Jars list needed to run
  the bundle. It lists the mongoDB Java driver as one of the required
  dependencies deployed in the same package;
  \item \textbf{Import-Package}: the Java Packages needed by the DSP Data
  Persistence Component Bundle. These Java Packages must be provided by other
  DSP components, forcing the execution to be dependent on those packages.
  They are provided through a section called ``Export-Package'', in other
  OSGi bundles and as it is shown in Listing \ref{file:OSGi-manifest}, they
  are can be from different packages.
\end{itemize}

@@

\begin{itemize}
  \item \textbf{config.xml}: the DSP Data Persistence is added to the DSP
  Platform by adding an entry for the component's OSGi Bundle artifact. This
  artifact can be seen in Listing \ref{file:dsp-config.xml};
  \item \textbf{matcher\underline{ }config.xml}: the addition of the rules that
  filters DSP Messages to the DSP Data Persistence component;
  \item \textbf{DSPDataPersistence\underline{ }001.xml}: depicts the bootstrap
  message for the DSP Data Persistence Component, as shown in Listing 
  \ref{file:dsp-data-persistence-bootstrap.xml}.
\end{itemize}

The first configuration artifact enables NetBEAMS to initialize the
descriptor of the component, as shown in Listing \ref{file:dsp-config.xml}.
The component is added with the highest priority, since it contains
dependencies to other DSP Components.

@@

In order to start the server, the command-line shown in Listing
\ref{cmd:mongodb-server-start} is used, where the data is configured to be
stored at the directory  ``./data/'', relative to the directory shown in the
shell. Also, the port number where the server listens to the client
connections is give as port ``20000". Once the server is running, any client
application can access to the server. For example, in order to connect the DSP
Data Persistence component, the port 20000 is an optional value be provided in
the DSP Data Persistence bootstrap message, as shown in Listing
\ref{file:dsp-data-persistence-bootstrap.xml}. Although mongoDB offers the
option to run with its specific default port number, it is recommended to
assign different ports, once the use of shards requires different port numbers
for each of the shard servers. An example of accessing mongoDB using a shell
client is shown in Listing \ref{cmd:mongodb-client-start}.

@@

The iterative mongo client shell offers users to verify and navigate on a
given database and its collections. This first section shows the connection of
the mongo client to the database ``netbeams'' as on Listing
\ref{cmd:mongodb-server-start}. The execution of the operation to count the
existing number of documents in the mongoDB collection ``netbeams'' is shown in
Listing \ref{cmd:mongodb-count}. It highlights the query for the collections
available \cite{mongodb} in the netbeams database.

@@

An example about data retrieving of a random document from the collection
``netbeams'', ``findOne()'' is documented to provide this support as shown in
Listing \ref{cmd:mongo-findone}. The output is the actual instance of a
document, showing all the collected data in the key ``observation'', and the
provenance metadata in the keys ``sensor.location'' and ``sensor.time''.

@@

The query based on attributes can be done using the "dot" notation, as the user
can navigate through the JSON documents. Additionally, the use of internal
functions is provided by mongoDB on the result of others. The example in
Listing \ref{cmd:mongo-find-list} counts the number of documents with the key
"data.ph" equals to "5.64".

@@

The mongoDB server offers different drivers to access the data, as well as the
REST Web Services. In order to get the results using the REST Web Services
interdace, the request can be issued using a web browser. For instance, the
request to the documents using the HTTP Request Variables to filter by the
fields as shown in Listing \ref{cmd:mongo-rest-request}.

>>>>

NetBEAMS uses the concept of the modularization through the use of the
Producer-Consumer paradigm described in the previous section. Since the DSP
Components are essentially OSGi bundles, the interoperability between
them is given by the OSGi Framework, described by an artifact called
MANIFEST.MF (Listing \ref{file:OSGi-manifest} on page
\pageref{file:OSGi-manifest}). In general, an OSGi
bundle must provide specifications that describe the module to be published into the OSGi
Framework. In this way, the main properties of the OSGi MANIFEST.MF artifact
can used by the DSP Data Persistence can be summarized as follows:

\begin{itemize}
  \item \textbf{Bundle-Activator}: The name of the instance of an OSGi
  Activator class, responsible to manage the bundle. The implemented class
  ``DataPersistenceActivator'' provides the activation mechanisms for the
  services of the DSP Data Persistence Component;
  \item \textbf{Bundle-ClassPath}: the necessary Java Jars list needed to run
  the bundle. It lists the mongoDB Java driver as one of the required
  dependencies deployed in the same package;
  \item \textbf{Import-Package}: the Java Packages needed by the DSP Data
  Persistence Component Bundle. These Java Packages must be provided by other
  DSP components, forcing the execution to be dependent on those packages.
  They are provided through a section called ``Export-Package'', in other
  OSGi bundles and as it is shown in Listing \ref{file:OSGi-manifest}, page
  \pageref{file:OSGi-manifest}, they are can be from different packages.
\end{itemize}

@@

\begin{itemize}
  \item \textbf{config.xml}: the DSP Data Persistence is added to the DSP
  Platform by adding an entry for the component's OSGi Bundle artifact. This
  artifact can be seen in Listing \ref{file:dsp-config.xml}, page
  \pageref{file:dsp-config.xml};
  \item \textbf{matcher\underline{ }config.xml}: the addition of the rules that
  filters DSP Messages to the DSP Data Persistence component;
  \item \textbf{DSPDataPersistence\underline{ }001.xml}: depicts the bootstrap
  message for the DSP Data Persistence Component, as shown in Listing 
  \ref{file:dsp-data-persistence-bootstrap.xml}.
\end{itemize}

The first configuration artifact enables NetBEAMS to initialize the
descriptor of the component, as shown in Listing \ref{file:dsp-config.xml},
page \pageref{file:dsp-config.xml}. The component is added with the highest
priority, since it contains dependencies to other DSP Components.

@@

In order to start the server, the command-line shown in Listing
\ref{cmd:mongodb-server-start} is used, where the data is configured to be
stored at the directory  ``./data/'', relative to the directory shown in the
shell. Also, the port number where the server listens to the client
connections is give as port ``20000". Once the server is running, any client
application can access to the server. For example, in order to connect the DSP
Data Persistence component, the port 20000 is an optional value be provided in
the DSP Data Persistence bootstrap message, as shown in Listing
\ref{file:dsp-data-persistence-bootstrap.xml}. Although mongoDB offers the
option to run with its specific default port number, it is recommended to
assign different ports, once the use of shards requires different port numbers
for each of the shard servers. An example of accessing mongoDB using a shell
client is shown in Listing \ref{cmd:mongodb-client-start} page
\pageref{cmd:mongodb-client-start}.

@@

The iterative mongo client shell offers users to verify and navigate on a
given database and its collections. This first section shows the connection of
the mongo client to the database ``netbeams'' as on Listing
\ref{cmd:mongodb-server-start}. The execution of the operation to count the
existing number of documents in the mongoDB collection ``netbeams'' is shown in
Listing \ref{cmd:mongodb-count}, page \pageref{cmd:mongodb-count}. It
highlights the query for the collections available in the netbeams database.

@@

An example about data retrieving of a random document from the collection
``netbeams'', ``findOne()'' is documented to provide this support as shown in
Listing \ref{cmd:mongo-findone}, page \pageref{cmd:mongo-findone}. The output
is the actual instance of a document, showing all the collected data in the
key ``observation'', and the provenance metadata in the keys
``sensor.location'' and ``sensor.time''.

@@

The query based on attributes can be done using the "dot" notation, as the user
can navigate through the JSON documents. Additionally, the use of internal
functions is provided by mongoDB on the result of others. The example in
Listing \ref{cmd:mongo-find-list}, page \pageref{cmd:mongo-find-list} counts
the number of documents with the key "data.ph" equals to "5.64".

@@

The mongoDB server offers different drivers to access the data, as well as the
REST Web Services. In order to get the results using the REST Web Services
interdace, the request can be issued using a web browser. For instance, the
request to the documents using the HTTP Request Variables to filter by the
fields as shown in Listing \ref{cmd:mongo-rest-request}, page
\pageref{cmd:mongo-rest-request}.

### 2009-12-14 docs/marcello-dissertation/thesis/chapter9-conclusions-and-future-works.tex
Final Revision After Prof. Puder's Audio Feedback. Chapters 7 and Chapter 8 were included together. Chapter 10 also added as the appendix, instead of chapter.
* Chapter 7 

- The mouse experiment was used as demo purpose and to show the DSP use of the persistence. The data used was the for the workload used the YSI Generator from Teresa, which generated the SAME values each time… In this way, it was NOT random… I added the proper random functions in everything, including the parameter passed to the method to generate a given number of generated values. 

- The workload uses the amount of data

There were significant remarks covered in this work, including the
importance that sensor networks are becoming ubiquitous, having their produced
data available online and other types of media. In order to provide access to
the collected data, one must first assess the different types of
infrastructural characteristics of the studied sensor networks. The selection
of a technology that provides such a persistence solution capabilities can be 
part of a process of analysis of taxonomies proposed by this work,
characterized as the first contribution of this report.

Based on the findings of the taxonomies, together with the empirical analysis
of the technologies, this work revealed a novel approach to provide
persistence and access to collected data from sensor devices using the
Key-Value-Pair data model. Besides the contribution of a new component for the
Data Sensor Platform, the evidence that the system improves productivity was
confirmed during the execution of the experiments, writing less source-code to
obtain better functionalities when compared to other solutions. Similarly,
improved work suggests that the abstraction of programming language should be
used with research groups whose primary expertise is not in Computer Science
or Database Systems.

The DSP Data Persistence component improves NetBEAMS's support to SF-BEAMS,
incorporating a persistence layer for the sensor network that can be used in
different ways. The solution for data persistence using Key-Value Pair
databases is a novel way to save collected data from sensor devices into a
single or multiple servers of a database technology. mongoDB's ability to cope
with the uncertainty of any properties collected from any sensor specification
is the most compeling reason to use the Key-value pair data model for collected
data from sensor devices.

As far as future works, there are many different directions for exploration.
The first major release of mongoDB with complete support to shards is
scheduled for 2010. An important mechanism of mongoDB is that it can
manage different mongoDB servers spread out in a cluster. This feature will give
better support to the Data-Centric Storage approach. This approach can reduce
the retrieval time by focusing on centralized selection of data over a given
database shard. Moreover, the overall search through the collected data
attributes can be drastically decreased by a parallel search over the database
shards using the MapReduce \cite{map-reduce} tools. At the time of development
of the experiments, the MapReduce API from mongoDB was also in the alpha
version, and therefore, presented many bugs. One suggestion for the
implementation of such functionality is based in Figure
\ref{fig:future-works-data-centric-map-reduce}, which shows an architectural
idea of the deployment of a Data-Centric storage approach. When searching for
specific document attributes, the process is sent in parallel to each of the
database shards, and the cluster head collects the result. After calculating
the final result, mongoDB server returns the relevant result to the user.

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.65]{../diagrams/future-works-data-centric-map-reduce}
  \caption{Data-Centric approach using MapReduce to search across different
  database shards}
  \label{fig:future-works-data-centric-map-reduce}
\end{figure}

In addition to the infrastructure, the use techniques for scheduling data
gathering from the SF-BEAMS sensor networks can be used to decrease the data
load into the system. While the period of times used by RTC seems
to cover the basic requirements of data, the data storage may be
seen as overloaded with repeated data. Since environmental conditions
might not change drastically over time, the use of data scheduling and
data clustering approaches can be used. Considering that SF-BEAMS
contains a cluster of sensors, the data load could be decreased by using
techniques such as clustering the data in the data sink before they are
persisted \cite{sn-time-series, sn-schedule-minimal-aggregation-time}. Since
the infrastructure of the SF-BEAMS sensor network is a one-hop start design,
the knowledge about the data can only be achieved in the data sink. One
suggestion is the implementation of a DSP Data Clustering component that is
responsible for the data clustering before they are inserted into the
database. Similarly, the use of schedulers on the sensors could also be used
to decrease the amount of data to be sent to the data sink, as seen at the
survey \cite{sn-scheduling-survey}.

>>>>

There were significant trends covered in this work, including the importance
that sensor networks are becoming ubiquitous, having their produced data
available online and other types of media. In the scope of Environmental Sensor
Networks, the collected data is the most important parameter to real-world
applications based on sensor networks. For example, the data collected in the
water may determine if a river is polluted, while the data collected during a
heavy snowfall can indicate hazard alerts the population. However, in order to
provide access to the collected data, one must first assess the different types of
infrastructural characteristics of the studied sensor networks. The selection
of a technology that provides such a persistence solution capabilities can be
part of a process of analysis of taxonomies proposed by this work,
characterized as the first contribution of this report.

Based on the findings of the taxonomies, together with the empirical analysis
of the technologies, this work revealed a novel approach to provide
persistence and access to collected data from sensor devices using the
Key-Value-Pair data model. Besides the contribution of a new component for the
Data Sensor Platform, the evidence that the system improves productivity was
confirmed during the execution of the experiments, writing less source-code to
obtain better functionalities when compared to other solutions. Similarly,
this work suggests that the abstraction of programming language should be
used with research groups whose primary expertise is not in Computer Science
or Database Systems.

The DSP Data Persistence component improves NetBEAMS's support to SF-BEAMS,
incorporating a persistence layer for the sensor network that can be used in
different ways. The solution for data persistence using Key-Value Pair
databases is a novel way to save collected data from sensor devices into a
single or multiple servers of a database technology. mongoDB's ability to cope
with the uncertainty of any properties collected from any sensor specification
is the most compelling reason to use the Key-value pair data model for collected
data from sensor devices.

As far as future works, there are many different directions for exploration.
The first major release of mongoDB with complete support to shards is
scheduled for 2010. An important mechanism of mongoDB is that it can
manage different mongoDB servers spread out in a cluster. This feature will give
better support to the Data-Centric Storage approach. This approach can reduce
the retrieval time by focusing on centralized selection of data over a given
database shard. Moreover, the overall search through the collected data
attributes can be drastically decreased by a parallel search over the database
shards using the MapReduce \cite{map-reduce} tools. At the time of development
of the experiments, the MapReduce API from mongoDB was also in the alpha
version, and therefore, presented many bugs. One suggestion for the
implementation of such functionality is based in Figure
\ref{fig:future-works-data-centric-map-reduce} (\cite{map-reduce-notes}), which
shows an architectural idea of the deployment of a Data-Centric storage approach. One scenario of
data retrieval is when a given processing in the search results are
required to be executed in parallel. The task of querying broken down into
two phases: the definition of the reduced problem as the parts, and the
mapping of the final result calculated with the response from each part. If
``Task" is considered to be a query (arrow 1) tor the server, then the
execution of the same query occurs in all different partitions (arrow 2), or
mongoDB shards. After processing the request, each node returns the result of
the query to the server (arrow 3), whose responsibility is to process each
individual result and return the final Result to the client application (arrow
4).

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.65]{../diagrams/future-works-data-centric-map-reduce}
  \caption{Data-Centric approach using MapReduce (\cite{map-reduce-notes})}
  \label{fig:future-works-data-centric-map-reduce}
\end{figure}

In addition to the infrastructure, the use  scheduling techniques for
data gathering in sensor networks can help the SF-BEAMS sensor networks
decrease the data load into the system. While the period of times used by RTC
seems to cover the basic requirements of data, the data storage may be seen as
overloaded with repeated data. Since environmental conditions might not change
drastically over time, the use of data scheduling and data clustering
approaches can be used. Considering that SF-BEAMS contains a cluster of
sensors, the data load could be decreased by using techniques such as
clustering the data in the data sink before they are persisted
\cite{sn-time-series, sn-schedule-minimal-aggregation-time}. Since the
infrastructure of the SF-BEAMS sensor network is a one-hop start design, the
knowledge about the data can only be achieved in the data sink. One suggestion
is the implementation of a DSP Data Clustering component that is responsible
for the data clustering before they are inserted into the database. Similarly,
the use of schedulers on the sensors could also be used to decrease the amount
of data to be sent to the data sink, as seen at the survey
\cite{sn-scheduling-survey}.
